The Resurgence of the Linear Optics Interferometer --- Recent Advances \& Applications

### 2017-04-17 paper.tex
Added more to characterization section.

\comment{Discuss the Reck et al. decomposition}

@@

\section{Reconstructing the linear optical network}
\sihui{Discuss the role of quantum tomography here. 
Algorithms using O'Brien and Laing, Schaffner and Tillmann--using single-photon, and two-photon inputs to reconstruct the Euler angles of the circuit.
Others using coherent-state inputs which has a lower requirement on the state-preparation side.}

@@

Generally, the unitary matrix of the $d\times d$ linear optical device are complex numbers $U_{ij}=r_{ij}e^{i\theta_{ij}}$, where $0 \leq r_{ij}\leq 1$, and $0\leq \theta_{ij}\leq 2\pi$. The scheme in \cite{bib:Laing12} relies on injecting one-photon and two-photon states into the linear optical network with correlated photon detection. First, they note some equivalencies: two unitaries $U$ and $U'$ are equivalent if there exist two diagonal unitary matrices $D_1^U$ and $D_2^U$ such that $U'=D^U_1 U D^U_2$, because these diagonal matrices are regarded as unknown and trivial phases on the input and output ports of the network, to which the one-photon and two-photon data are insensitive to. This reduces the first row and first column elements to real numbers, i.e. $\theta_{1j}=\theta_{i1}=0$. Second, the photon statistics remain unchanged under the complex conjugation of $U$. Thus, the imaginary part of $M_{2,2}$ must be non-negative.

>>>>

@@

\section{Reconstructing the linear optical network}

@@

Generally, the unitary matrix of the $d\times d$ linear optical device are complex numbers $U_{ij}=r_{ij}e^{i\theta_{ij}}$, where $0 \leq r_{ij}\leq 1$, and $0\leq \theta_{ij}\leq 2\pi$. The scheme in \cite{bib:Laing12} relies on injecting one-photon and two-photon states into the linear optical network with correlated photon detection. First, they note some equivalencies: two unitaries $U$ and $U'$ are equivalent if there exist two diagonal unitary matrices $D_1^U$ and $D_2^U$ such that $U'=D^U_1 U D^U_2$, because these diagonal matrices are regarded as unknown and trivial phases on the input and output ports of the network, to which the one-photon and two-photon data are insensitive to. This reduces the first row and first column elements to real numbers, i.e. $\theta_{1j}=\theta_{i1}=0$. Second, the photon statistics remain unchanged under the complex conjugation of $U$. Thus, the imaginary part of $M_{2,2}$ must be non-negative. Then assuming that the first two rows and columns are non-vanishing, and that there is no total loss in the interferometer, the matrix to be recovered is
\begin{align}
U=\left (\begin{array}{cccc}
r_{11} & r_{12} & \ldots & r_{1m} \\
r_{21} & r_{22}e^{i\theta_{22}} &\ldots & r_{2m}e^{i\theta_{2m}}\\
\vdots & \vdots & \ddots & \vdots \\
r_{m1} & r_{m2}e^{i\theta_{m2}} & \ldots & r_{mm}e^{i\theta_{mm}}
 \end{array}\right ) \ .
\end{align}
In the paper, they showed that it is possible to write the paramaters $r_{ij}$, and $\theta_{ij}$ for $i,j\geq 2$ in terms of the $2m-1$ real parameters of the first column and row, and the visibility of two-photon inputs. The remaining $2m-1$ real parameters can be found via one-photon 
transmissions.  An increased accuracy of the characterization is possible by estimating and correcting systematic errors that arise due to mode mismatch \cite{Dhand16}. Others have used numerical methods to find the closest parameters that yield the observed visibilities \cite{bib:Spagnolo16,bib:Tillmann16}.

Another characterization method was presented that is similar to \cite{bib:Laing12} with the important exception that coherent states are to be used instead \cite{bib:Rahimi-Keshari13,bib:Heilmann15}. Such states are produced by a standard laser source, thus reducing the resource needed. The $r_{jk}$ terms are found by the square root of the ratios of output intensity at the $k$th port to the input intensity at the $j$th port. The remaining phases $\theta_{ij}$ are found by the interference pattern given by a two-mode coherent state $\ket{\alpha_1}\ket{\alpha_2}$ created by splitting a single coherent state on a 50-50 beamsplitter, and then imparting a relative phase, $\phi$, between them. The states $\ket{\alpha_1}$ and $\ket{\alpha_2}$ are input into port 1, and $j$ respectively. The output intensity at the $j$th port is 
\begin{align}
I_k=I(r_{1k}^2+r_{jk}^2+2 r_{1k}r_{kj}\cos(\phi+\theta_{jk})) \ ,
\end{align}
where $\theta_{jk}=0$ for $k=1$. By scanning the phase shift $\phi$ and then locating the maximum value of $I_k$ for $j=2,\ldots, m$, all unknown phases can be found via $\theta_{jk}=2\pi-\phi$. An elegant extension of the scheme of Rahimi-Keshari {\it et al.} removes the need for precise control of the phase shift $\phi$ \cite{bib:Heilmann15} by suggesting instead to plot the output intensity $I_k$ with respect to the input intensity $I$. In time, the natural drift in the laser source will cause this plot to trace out an ellipse, known as a Lissajous figure. whose orientation and direction of evolution will give the phase $\theta_{jk}$ and its sign respectively.
